Robust and Efficient CSI Compression Using CLLWCsiNet in Noisy MIMO Environments

### 2025-04-07 paper_latex_edit - Copy.tex
correct reference 25
Litrature-review-sort-by-year-in-citation

For the first time, Wen et al. \cite{abe} introduced a deep learning solution for CSI feedback compression by employing an auto-encoder architecture. Their work serves as the foundation for numerous subsequent research efforts aiming to incorporate the behavior of wireless communication systems within MIMO systems using an auto-encoder architecture. In this setup, the encoder component assumes the role of the user equipment, while the decoder portion functions as the decoder. The random spatial characteristics of channel state information have been simulated using the COST2100 channel model \cite{abf}. The COST2100 dataset has been generated accordingly. Proposed architectures have been predominantly founded upon auto-encoder structures. Wen et al. \cite{abe} introduced CsiNet, an autoencoder that employs the Residual Network \cite{abg} concept in the decoder part. Wang et al \cite{abh}. introduced time-varying CSI Feedback, aiming to simulate the coherence time in the wireless channel attributed to user equipment mobility. They incorporated Long Short-Term Memory (LSTM) units into the decoder part of an auto-encoder. Lu et al. \cite{abi} implemented the LSTM architecture in both the encoder and decoder parts of their model. They opted for a neural network instead of a convolutional form for LSTM, which resulted in a doubling of parameters and complexity compared to CsiNet-LSTM \cite{abh}, as reported. They emphasized memory for feature extraction module more than correlation by employing LSTM in parallel with a fully connected layer, akin to jump connections in residual networks \cite{abi}. Their proposed architecture, RecCsiNet \cite{abi}, achieved superior results prior to considering time correlation, by employing LSTM for residual features instead of correlation features \cite{abh}. Li and Wu \cite{abj} extended the RecCsiNet model by separating the feature extraction process from feature compression and decompression. Their proposed model, Convlstm, significantly reduces the error by a considerable amount \cite{abj}. Liu et al. \cite{abk} emphasized coherence time simulation through LSTM and proposed a new architecture called Markovenet. In addition to their suggested architecture and a simple Markov model, they introduced a new convolutional latent space that significantly reduces the number of parameters and complexity of the autoencoder. They argue that instead of using fully connected layers to extract features, which may not be crucial in the random characteristics of the channel state information matrix, a convolutional approach is more appropriate, given that only adjacent elements in the channel state information matrix exhibit strong correlations. Mashhadi et al. \cite{abl} proposed the DeepCMC network, which is a fully convolutional neural network. Dimensional compression is achieved through upsampling and downsampling in pooling layers. To further compress the representation, they utilize a context-adaptive binary arithmetic coding scheme for the encoder side of the DeepCMC output codeword. Liang et al. \cite{abm} proposed a compressive sensing architecture named CSRenet, which employs compressive sensing on the encoder side and a deep learning-based method on the decoder side. Lu et al. \cite{abn} introduced the CRNet architecture, representing one of the pioneering studies utilizing small kernel sizes for CSI matrix feature granularity extraction. They employed a residual network architecture in both the encoder and decoder. Within the residual architecture, Lu et al. utilized different kernel sizes in different paths and suggested a multi-resolution block for both encoder and decoder. An important aspect of their research is that the performance of a single-resolution refine network can be enhanced by employing different kernel sizes. Guo et al. \cite{abo} proposed a multiple-rate architecture to address the challenge posed by short coherence time intervals, where CSI must be sufficiently compressed and the auto-encoder should dynamically adjust the compression ratio according to the environment. Their proposed multiple-rate framework is capable of compressing the CSI matrix at different compression ratios without altering the architecture or requiring a new training scheme. As a result, user equipment (UE) does not need to store parameters for different architectures. Guo et al. incorporated quantization and dequantization modules in the encoder and decoder, respectively. In contrast to Lu et al. \cite{abp}, who used uniform quantization, Guo et al.\cite{abo} employed non-uniform quantization to generate binary streams at the output of the encoder. Additionally, they suggested using larger kernel sizes in both the encoder and decoder, as well as an expected residual configuration. These changes were applied to CsiNet, resulting in a new architecture called CsiNetPlus \cite{abo}. Yu et al. \cite{abq} employ the non-local block which enhance the accuracy with the help of long-range dependences. They inspired the non-local block from X et al. \cite{abr} which has been for sequence data denoising. Cai et al. \cite{abs} proposed a modified version of CsiNet called Attention-CsiNet. They replaced the fully connected layers in the encoder and decoder with LSTM modules to extract coherence among sub-carriers. Additionally, Cai et al. \cite{abs} introduced an Attention module for CSI feedback compression, inspired by the research of \cite{abt}. J et al. \cite{abt} proposed the squeeze-and-extraction network in the field of machine translation.Lu \textit{et al.} \cite{prac2024} identify key challenges including channel distribution shifts and propose an alternating optimization framework with online knowledge-review to enhance adaptability. Their solution improves NMSE by 4.3 dB in unseen environments compared to conventional static models.Recent work by Yin \textit{et al.} \cite{quant2024} proposes a novel bit allocation strategy for encoder outputs in DL-based CSI feedback, dynamically assigning more bits to high-variance features. Their alternating training method jointly optimizes quantization codebooks with autoencoder weights while using an adaptive logarithmic loss function to prevent gradient vanishing. This approach reduces NMSE by 2.1 dB compared to uniform quantization at 4-bit compression.
Ma \textit{et al.} \cite{Ma2024} present a novel learning-to-optimize (L2O) framework that combines compressive sensing with deep learning for efficient CSI feedback. Their Csi-L2O approach achieves state-of-the-art performance (-18.88 dB NMSE) while reducing encoder complexity to just 0.15\% of conventional methods through innovative linear projection and LSTM-based optimization.
\cite{CSICompression2024} systematically analyzes CSI feedback overhead challenges in massive MIMO, demonstrating that uncompressed feedback can reach 2 Mb/s per antenna. The study provides a comprehensive comparison between traditional model-driven compression and emerging data-driven approaches, noting that while machine learning methods achieve better accuracy, they incur higher computational costs. The authors evaluate deployment readiness for 5G-Advanced systems, highlighting critical trade-offs between efficiency and flexibility in practical implementations.
\cite{Guo2024} addresses two critical challenges in DL-based CSI feedback: vendor coordination overhead and modular design inefficiencies. The authors propose a \textit{one-sided} CSI-PPPNet architecture requiring only 4.3\% of conventional model parameters (175K vs. 4095K), while supporting variable compression ratios without retraining. Their joint multi-module learning framework demonstrates a 15--20\% spectral efficiency improvement over decoupled approaches by simultaneously optimizing CSI feedback, channel coding, and precoding modules.
Shehzad et al. \cite{shehzad2021design} proposes a novel machine learning-based approach for efficient Channel State Information (CSI) feedback in massive MIMO systems. The authors address the critical challenge of feedback overhead reduction while maintaining CSI accuracy, which is essential for realizing the full potential of massive MIMO technology. Their key innovation is a hybrid approach that combines conventional CSI feedback with twin recurrent neural network (RNN) predictors at both the base station (gNB) and user equipment (UE).
Zhang et al. \cite{zhang2022quantization} addresses the critical challenge of quantization in deep learning-based CSI feedback for massive MIMO systems. The authors propose an innovative adaptor-assisted quantization strategy that significantly improves feedback efficiency while maintaining reconstruction accuracy.
\cite{liu2023continuous} tackles a critical limitation of deep learning (DL)-based CSI feedback in massive MIMO systems: the performance degradation caused by distribution shifts between training and real-world deployment scenarios. The authors propose a continuous online learning framework that addresses catastrophic forgetting while maintaining high feedback accuracy across diverse channel conditions.
\cite{wang2024better} introduces a novel test channel-based quantization module (TCQM) that addresses critical limitations in DL-based CSI feedback systems.
\cite{zhang2024zone} proposes a novel zone-specific deep learning approach for CSI feedback in massive MIMO systems. The key innovation partitions the coverage area into spatial zones with distinct channel characteristics, training specialized autoencoders for each zone.
\cite{chen2024transnet} propose TransNet+, an enhanced Transformer-based architecture for massive MIMO CSI feedback. The method improves upon existing Transformer designs by incorporating higher-dimensional feature representations and an optimized multi-head attention mechanism. A warm-up-aided cosine training scheme is introduced to accelerate convergence and enhance feature learning. Experimental results demonstrate superior performance compared to state-of-the-art methods, particularly at high compression ratios, with significant NMSE reduction in both indoor and outdoor scenarios.
\cite{li2024lightweight}Proposes GCRNet, a Ghost module-based architecture that reduces feature redundancy in CSI feedback by generating secondary features from primary maps. Achieves 22.15\% lower complexity than \cite{ji2021clnet} while maintaining comparable accuracy, addressing efficiency decay in convolution-based methods like \cite{abe}.
\cite{wang2024variational}Proposes a variational autoencoder framework for noisy-channel CSI feedback, formulating it as rate-distortion optimization. The method achieves 40\% bandwidth reduction versus \cite{abe} while maintaining NMSE robustness across SNRs, addressing limitations of uniform bit treatment in \cite{abn,chen2024transnet}.
\cite{mashhadi2020deepcmc}Proposes DeepCMC, a fully convolutional autoencoder with quantization and entropy coding blocks that achieves 3--5~dB lower NMSE than \cite{abe} at equivalent bit rates. Extends to a distributed version exploiting inter-user CSI correlations, reducing feedback overhead by 17--50\% for nearby users (<1m spacing) compared to single-user compression.
In this paper, we address the inefficiencies associated with the fully connected layers in many state-of-the-art models for CSI feedback compression. These layers often propagate redundant weights related to the zero parts of the channel state information, leading to a large number of parameters. This inefficiency is particularly problematic on User Equipment (UE), where resource constraints such as memory and computational power are critical factors. By focusing on quantization techniques, we aim to significantly reduce the parameter count and enhance the efficiency of CSI feedback compression models on UE.

@@

	\bibitem{CSICompression2024}
	A. Author et al., 
	"Title of the Paper in Sentence Case," 
	\textit{IEEE Journal Name}, 
	vol. XX, no. X, 
	pp. XXX--XXX, 
	Month. Year, 
	doi: XXXXX.XXXXXX.

@@

>>>>

Pioneering the application of deep learning to CSI feedback compression, Wen et al. \cite{abe} introduced an auto-encoder architecture. This seminal work laid the groundwork for extensive subsequent research that explored the integration of wireless communication system dynamics within MIMO systems using auto-encoder frameworks. In this initial setup, the encoder emulated the user equipment, while the decoder functioned at the receiver. The inherent randomness of channel state information was simulated using the COST2100 channel model \cite{abf}, leading to the creation of the corresponding COST2100 dataset.

Building upon this foundation, numerous architectures, predominantly based on auto-encoders, have been proposed. Wen et al. further developed this line of research with CsiNet \cite{abe}, an autoencoder notable for incorporating the Residual Network concept \cite{abg} within its decoder. Addressing the temporal aspects of wireless channels affected by user mobility, Wang et al. \cite{abh} introduced time-varying CSI Feedback by integrating Long Short-Term Memory (LSTM) units into the decoder of an auto-encoder. Lu et al. \cite{abi} expanded on this by implementing LSTM in both the encoder and decoder of their model. However, their choice of a fully connected network for LSTM, instead of a convolutional structure, reportedly doubled the parameters and complexity compared to CsiNet-LSTM \cite{abh}. Their proposed RecCsiNet \cite{abi} prioritized memory for feature extraction over direct correlation, employing LSTM in parallel with a fully connected layer, similar to skip connections in residual networks \cite{abi}, achieving superior performance before considering temporal correlation by applying LSTM to residual features rather than correlation features \cite{abh}.

Further refining these approaches, Mashhadi et al. \cite{mashhadi2020deepcmc} introduced DeepCMC, a fully convolutional neural network where dimensional compression is achieved through pooling layers. To enhance compression, they also employed context-adaptive binary arithmetic coding on the encoder's output. Shehzad et al. \cite{shehzad2021design} proposed a novel machine learning-based approach utilizing twin recurrent neural networks at both the base station and user equipment to reduce CSI feedback overhead while maintaining accuracy. Zhang et al. \cite{zhang2022quantization} tackled the critical issue of quantization in DL-based CSI feedback by proposing an adaptor-assisted quantization strategy to improve feedback efficiency and accuracy. Liu et al. \cite{liu2023continuous} addressed the challenge of performance degradation due to distribution shifts by proposing a continuous online learning framework to maintain accuracy across diverse channel conditions. Guo et al. \cite{guo2023ris} introduced RIS-CoCsiNet, a cooperative CSI feedback framework for RIS-assisted systems, separating RIS-UE CSI into shared and unique components and achieving significant NMSE improvements through magnitude-dependent phase feedback strategies.

More recent advancements continue to push the boundaries of DL-based CSI feedback. Li and Wu \cite{abj} extended the RecCsiNet model by decoupling feature extraction from compression and decompression, resulting in their Convlstm model, which reportedly significantly reduced error \cite{abj}. Liu et al. \cite{abk} also focused on coherence time simulation with LSTM and introduced Markovenet, featuring a novel convolutional latent space that reduces model parameters and complexity. They argued for the superiority of convolutional approaches over fully connected layers for feature extraction in CSI matrices due to the localized correlations. Lu et al. \cite{abn} presented CRNet, a significant early work employing small kernel sizes for fine-grained feature extraction and utilizing a residual network in both encoder and decoder with multi-resolution blocks. Guo et al. \cite{abo} proposed a multiple-rate architecture (CsiNetPlus), incorporating non-uniform quantization and larger kernel sizes to dynamically adjust compression ratios based on the environment without retraining. Yu et al. \cite{abq} explored the use of non-local blocks, drawing inspiration from sequence denoising work \cite{abr}, to capture long-range dependencies and enhance accuracy. Cai et al. \cite{abs} introduced Attention-CsiNet, replacing fully connected layers with LSTMs for sub-carrier coherence extraction and incorporating an attention module inspired by machine translation research \cite{abt}. Liang et al. \cite{abm} proposed CSRenet, a compressive sensing architecture combining compressive sensing at the encoder with deep learning at the decoder.

Continuing this trend of innovation, recent studies in 2024 have further refined DL-based CSI feedback techniques. Wang et al. \cite{wang2024better} introduced a novel test channel-based quantization module (TCQM) to address limitations in existing systems. Zhang et al. \cite{zhang2024zone} proposed a zone-specific deep learning approach, training specialized autoencoders for different spatial zones to account for varying channel characteristics. Chen et al. \cite{chen2024transnet} presented TransNet+, an enhanced Transformer-based architecture leveraging higher-dimensional features and an optimized multi-head attention mechanism, achieving superior performance, especially at high compression ratios. Li et al. \cite{li2024lightweight} proposed GCRNet, a Ghost module-based architecture that reduces feature redundancy, achieving comparable accuracy to prior work with significantly lower complexity \cite{ji2021clnet}. Wang et al. \cite{wang2024variational} introduced a variational autoencoder framework for noisy channels, optimizing rate-distortion and demonstrating robustness across different SNRs. Ma et al. \cite{Ma2024} presented Csi-L2O, a learning-to-optimize framework combining compressive sensing and deep learning, achieving state-of-the-art performance with significantly reduced encoder complexity. Lu et al. \cite{prac2024} identified challenges like channel distribution shifts and proposed an alternating optimization framework with online knowledge review to enhance adaptability, demonstrating substantial NMSE improvements in unseen environments. Yin et al. \cite{quant2024} proposed a dynamic bit allocation strategy for encoder outputs, jointly optimizing quantization codebooks and autoencoder weights to reduce NMSE. \cite{CSICompression2024} provided a comprehensive analysis of CSI feedback overhead and a comparison of traditional and data-driven approaches, highlighting trade-offs for 5G-Advanced deployment. Finally, \cite{Guo2024} introduced a one-sided CSI-PPPNet architecture with a joint multi-module learning framework, significantly reducing parameters and improving spectral efficiency by simultaneously optimizing feedback, coding, and precoding modules.
In this paper, we address the inefficiencies associated with the fully connected layers in many state-of-the-art models for CSI feedback compression. These layers often propagate redundant weights related to the zero parts of the channel state information, leading to a large number of parameters. This inefficiency is particularly problematic on User Equipment (UE), where resource constraints such as memory and computational power are critical factors. By focusing on quantization techniques, we aim to significantly reduce the parameter count and enhance the efficiency of CSI feedback compression models on UE.

@@

	\bibitem{CSICompression2024}
	J.~Chen, W.~J.~Hillery, K.~R.~Mestav, C.~Nuzman, I.~Saniee, and Y.~Xing, 
	``CSI Compression for Massive MIMO: Model-Based or Data-Driven?,'' 
	\emph{IEEE Wireless Communications}, vol.~32, no.~1, pp.~22--27, Feb. 2025,

@@

	\bibitem{guo2023ris}
	 J. Guo, X. Yang, C.-K. Wen, S. Jin, and G. Y. Li, ``Deep learning-based CSI feedback for RIS-assisted multi-user systems,'' \textit{arXiv preprint arXiv:2003.03303}, 2023.

	 \bibitem{guo2023ris}
	 J. Guo, X. Yang, C.-K. Wen, S. Jin, and G. Y. Li, ``Deep learning-based CSI feedback for RIS-assisted multi-user systems,'' \textit{arXiv preprint arXiv:2003.03303}, 2023.
